---
title: "Brownlows Bets"
author: "Cleo Ware"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Fall 2023"    
    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
source(file="BrLw.R")
```


<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.


## Our strategy is to build a Function and the use it to answer the three questions. 

The function will be given a class_size by number of students, odds, and Number_of_times brownlow made the bet.  It will give you back how much Brownlows earnings were. We will use it to answer our three questions below. If you can, improve this function by giving default values to some of its inputs. 

# In the class and payoff as described  If you highered 40 brownlow substitute teachers to do the same thing that Brownlow did then.




## what is their average earnings & the standard deviation and a 95% confidence interval for their earnings?

```{r}
options(digits = 2)
teachers <- numeric(40)
for (i in 1:40){
  teachers[i]<-BrLw()
}

fav_stats(teachers) -> profit

print("we can say that Mr. Brownlows is making an average of")
print(profit$mean)
print("The 95% confidence level goes from")
print(profit$mean-2*profit$sd)
print("to")
print(profit$mean+2*profit$sd)

```


```{r}
central <- favstats(teachers)

cat("Mr Brownlow is on average making $",central$mean," over the course of the experiemnt.\n")

cat("95% Conficence Interval: Low:$",central$mean - 2*central$sd,"- High $", central$mean + 2*central$sd)
```


## We can say that Mr. Brownlows is making an average of `r profit$mean` with a standard deviation of `r profit$sd`. 




## the standard deviation and a 95% confidence interval for their earnings?
  
<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows winnings (or losings)-->


# What are the fair odds? Use manipulation of odds in the function to tune Brownlows earnings to zero. 

```{r}
teachers <- numeric (1000)
 for (i in 1:1000){
   teachers [i] = BrLw(odds=4.39)
}
 favstats(teachers)

```

## To start, we tested for best odds with 40 teachers. After testing thousands of times, we were unable to find a stable mean. So, we added more teachers, to a total of 1000. After adding more teachers, we tested the odds again.After testing multiple sets of odds, we found that 4.39 was the most stable set of odds.


```{r}
# Create a histogram
hist(teachers, 
     main = "teachers Histogram",
     xlab = "teachers",
     ylab = "Frequency",
     col = "skyblue",
     border = "black")
```


<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 637 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->




# What size of class makes the 2:1 bet fair?  Use manipulation of number of Students to tune Brownlows earnings to zero.


```{r}
teachers <- numeric (40)
 for (i in 1:40){
   teachers [i] = BrLw(numClass= 28)
}
 favstats(teachers)

```


## After starting with 1000 students and trying different numbers, we decided to lower the number of students to 40. Still not able to get Brownlow's earnings down to zero, we tried 28 stduents. After seeing the results, we settled on 28 students because it gave us the closest number to zero. 


<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting profits!-->








# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->


## As we moved through this write up, we were able to manipulate the odds for both Brownlow and the students through trial and error before getting the desired results. We changed the number of teachers multiple times before settling 1000 to achieve the target mean. We were able to achieve this without any helper files. Although, we did receive the help of ChatGPT. 
